---
title: "Homework 4 - https://github.com/Henryblake2777/Homework3"
author: "Henry Blake - hrb987"
date: "2025-02-17"
output:
  pdf_document:
    toc: True
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      eval = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.align = "center")

library(readr)
library(ggplot2)
library(tidyverse)
library(rvest)
library(mosaic)

letter_freq <- read_csv('letter_frequencies.csv')
```

# Problem 1 - Iron Bank

## Null Hypothesis - 

H0 -> P = .024

## Test Statistic -

TS = 0.0346 - 0.024 = 0.0106

## P-Value -

```{r}
sim_trades <- do(100000)*nflip(n = 2001, prob = .024)
sum(sim_trades >= 70)/100000

```

## Probability Distribution -

```{r}
ggplot(sim_trades) + geom_histogram(aes(x = nflip), color = 'black', fill = '#B9375E') + labs(
  title = 'Distribution of Trials Over 70 Flagged Trails'
)
```

## Conclusion -

Due to a p-value of .0013, there is significant evidence that the 70 flagged by the detection algorithm were unusually high and not due to simple variability.

# Problem 2 - Health Inspections - 

## Null Hypothesis - 

H0 -> P = .03

## Test Statistic -

TS  = .16 - .03 = .13

## P-Value -

```{r}
sim_health <- do(100000)*nflip(n = 50, prob = .03)
sum(sim_trades >= 8)/100000

```

## Probability Distribution -

```{r}
ggplot(sim_health) + geom_histogram(aes(x = nflip), color = 'black', fill = '#BE9A60',binwidth = 1) + labs(
  title = 'Distribution of Trials Over 8 Failed Inspections',
  x = 'Proportion'
)
```

## Conclusion -

Due to a p-value of practically 0, there is very significant evidence that the 8 failed health inspections by Gourmet Bites are not due to normal factors.

# Problem 3 - Jury Selection

## Null Hypothesis - 

H0 -> Observed frequencies are consistent with the model

## Test Statistic -

Ts = chi2 = 12.43

## Chi-Squared Value -

```{r}
expected <- c(Group1 = .3*240, Group2 = .25*240, Group3 = .2*240, Group4 = .15*240, Group5 = .1*240)
observed <- c(Group1 = 85, Group2 = 56, Group3 = 59, Group4 = 27, Group5 = 13)

chi_squared_statistic = function(observed, expected) {
  sum((observed - expected)^2 / expected)
}

juror_chi2 <- chi_squared_statistic(observed, expected)

chi2_sim = do(100000)*{
  simulated_counts = rmultinom(1, 240, expected)
  this_chi2 = chi_squared_statistic(simulated_counts, expected)
  c(chi2 = this_chi2) # return a vector with names and values
}

sum(chi2_sim > juror_chi2)/100000
```

## Probability Distribution -

```{r}
ggplot(chi2_sim) + geom_histogram(aes(x = chi2), color = 'black', fill = '#CEDDBB',binwidth = 1) + labs(
  title = 'Distribution of Chi-Squared values',
  x = 'chi-squared'
)
```

## Conclusion -

Due to a p-value of .0143, there is intriguing evidence that the distribution of jurors selected by the judge does not follow the expected distribution. However, a confounding variable that may affect this could be the type of cases that the judge works.

# Problem 4 - LLM Watermarking

## Part A: Null Distribution - 

```{r}
sentences <- readLines('brown_sentences.txt')
clean_sentences = gsub("[^A-Za-z] ", "", sentences)
clean_sentences = gsub('\\.', "", clean_sentences)
clean_sentences = gsub(' ', "", clean_sentences)
clean_sentences = toupper(clean_sentences)
letter_chi2s <- c()
for (sentence in clean_sentences){
  all_letters <- LETTERS
  observed <- table(strsplit(sentence, '')[[1]])
  for (letter in all_letters){
    if (!(letter %in% names(observed))){
      observed[letter] <- 0
    }
  }
  observed <- observed[all_letters]
  expected <- letter_freq$Probability * nchar(sentence)
  names(expected) <- letter_freq$Letter
  letter_chi2 <- chi_squared_statistic(observed, expected)
  letter_chi2s <- c(letter_chi2s, letter_chi2)
}
ggplot() + geom_histogram(aes(x = letter_chi2s), color = 'black', fill = '#434343') + labs(
  title = 'Distribution of Chi-Squared values',
  x = 'chi-squared'
)
```

## Part B: Checking For a Watermark - 

```{r}
test_sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)

# Example loop to print each sentence
clean_test_sentences = gsub("[^A-Za-z] ", "", test_sentences)
clean_test_sentences = gsub('\\.', "", clean_test_sentences)
clean_test_sentences = gsub(' ', "", clean_test_sentences)
clean_test_sentences = toupper(clean_test_sentences)
letter_Ps_test <- c()
for (sentence in clean_test_sentences){
  all_letters <- LETTERS
  observed <- table(strsplit(sentence, '')[[1]])
  for (letter in all_letters){
    if (!(letter %in% names(observed))){
      observed[letter] <- 0
    }
  }
  observed <- observed[all_letters]
  expected <- letter_freq$Probability * nchar(sentence)
  names(expected) <- letter_freq$Letter
  letter_chi2_test <- chi_squared_statistic(observed, expected)
  letter_P_test <- round(sum(letter_chi2s > letter_chi2_test)/length(letter_chi2s),3)
  letter_Ps_test <- c(letter_Ps_test, letter_P_test)
  
}
print(letter_Ps_test)
```

According to the p-values of the chi-squared statistics, sentence 6 seems to be the one that is watermarked because of the p-value of .009. It does not seem to follow the typical letter frequency distribution.